---
title: "4cast comparer"
output:
  flexdashboard::flex_dashboard:
    source_code: https://github.com/bcgov/4cast_viewer_comparison
    orientation: columns
runtime: shiny
resource_files:
- out/census.rds
- out/errors.rds
- out/lfs_means.rds
- out/new_vs_fcast.rds
- out/new_vs_old.rds
- out/prop_data.rds
---

```{r, include=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(conflicted)
library(htmlwidgets) #for function onRender
library(plotly)
conflicts_prefer(dplyr::filter)
options(scipen = 999)
#functions-----------------------
my_dt <- function(tbbl) {
  DT::datatable(tbbl,
                extensions = "Buttons",
                rownames = FALSE,
                options = list(
                  columnDefs = list(list(className = "dt-center", targets = "_all")),
                  paging = TRUE,
                  scrollX = TRUE,
                  scrollY = TRUE,
                  ordering = TRUE,
                  dom = "t",
                  pageLength = 64
                )
  )
}

get_rmse <- function(tbbl){
  sqrt(mean((tbbl$new-tbbl$original)^2))
}

get_smape <- function(tbbl){
  mean(abs(tbbl$new-tbbl$original)/(abs(tbbl$new)+abs(tbbl$original)))
}

get_mean <- function(tbbl){
  (mean(tbbl$new)+mean(tbbl$original))/2
}

```

```{r}
new_vs_old <- read_rds(here("out","new_vs_old.rds"))
files <- sort(unique(new_vs_old$which_file), decreasing = TRUE)
new_vs_fcast <- read_rds(here("out","new_vs_fcast.rds"))
errors <- read_rds(here("out","errors.rds"))
census <- read_rds(here("out", "census.rds"))
lfs_means <- read_rds(here("out","lfs_means.rds"))
prop_data <- read_rds(here("out","prop_data.rds"))
```

# Comparison with previous cut

## Inputs {.sidebar data-width="350"}

-  We are interested in identifying occupations where the Stokes current $C_t$ and previous $P_t$ cuts differ significantly both in absolute and relative terms. 

- The root mean square error measures absolute differences:

1)  $${\displaystyle \operatorname {RMSE} ={\sqrt {{\frac {1}{n}}\sum _{t=1}^{n}(C_{t}-P_{t})^{2}}}}$$

-  In contrast the symmetric mean absolute percent error measures relative differences.

2)  $${\displaystyle {\text{SMAPE}}={\frac {1}{n}}\sum _{t=1}^{n}{\frac {|C_{t}-P_{t}|}{|C_{t}|+|P_{t}|}}}$$

- Similar forecast (small absolute and relative differences) can be found in the bottom left, different forecasts (large absolute and relative differences) top right.   


```{r}
selectInput(
  "file",
  "Select an excel file to compare",
  files,
  selected = files[4]
)

which_file <- reactive({
  temp <- new_vs_old|>
    filter(which_file==input$file)|>
    pull(joined)
  temp[[1]]
})

renderUI({#need the render UI because the list of choices depends on the excel file
selectInput(
  "variable",
  "Select a variable",
  unique(which_file()$Variable),
  selected = unique(which_file()$Variable)[1]
)
})

renderUI({#need the render UI because the list of choices depends on the excel file
selectInput(
  "geographic_area",
  "Select a geographic area",
  unique(which_file()$`Geographic Area`),
  selected = unique(which_file()$`Geographic Area`)[1]
)
})

checkboxInput("origin", "Include the origin?", value = FALSE)

#reactive elements------------------------------

reactive_base <- reactive({
  if(length(unique(which_file()$Description))==1 & length(unique(which_file()$Industry))>1){
    temp <- which_file()|>
      filter(Industry!="All industries")|>
      rename(slice = Industry)|>
      select(-Description)
  }else{
    temp <- which_file()|>
      filter(Description!="Total")|>
      rename(slice = Description)|>
      select(-Industry)
  }
  temp|>
    filter(year>=year(today()),
           Variable==input$variable,
           `Geographic Area`==input$geographic_area)|>
    ungroup()|>
    select(-Variable, -`Geographic Area`, -NOC)|>
    group_by(slice)
})

reactive_tbbl <- reactive({
  reactive_base()|>
    mutate(year=as.numeric(year))|>
    filter(year >= max(year)-10)|>
    group_by(slice)|>
    nest()|>
    mutate(rmse=map_dbl(data, get_rmse),
           smape=map_dbl(data, get_smape),
           mean_value=map_dbl(data, get_mean)
          )|>
  arrange(desc(mean_value))|>
  select(-data)
})
```

## Column

### `r renderUI({paste(input$geographic_area, input$variable, sep=": ")})`

```{r}

js <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point', {text: point.text});
  });
}"

plotly::renderPlotly({
  req(input$variable)
  req(input$geographic_area)

  plt <- reactive_tbbl()|>
    ggplot(aes(rmse, 
               smape, 
               size=mean_value, 
               colour=mean_value, 
               alpha=mean_value, 
               text=slice)
           )+
  geom_point()+
  scale_x_continuous(trans="log10")+
  scale_y_continuous(trans="log10")+
  scale_colour_viridis_c(labels=scales::comma)+
  labs(x="Root Mean Squared Error", y="Symmetric Mean Absolute Percentage Error", colour=NULL)

  plotly::ggplotly(plt, tooltip = "text")|>
     plotly::config(displayModeBar = FALSE)|>
     onRender(js)
})
```

## Column

### `r renderUI({input$point$text})`

```{r}
plotly::renderPlotly({
  req(input$point$text)
  plt <- reactive_base()|>
    filter(slice==input$point$text)|>
    pivot_longer(cols = c("new","original"))|>
    group_by(name)|>
    mutate(year=as.numeric(year),
           cagr=(tail(value, n=1)/head(value, n=1))^(1/(max(year)-min(year))) - 1,
           )|>
    ggplot(aes(year, 
               value, 
               colour=name,
               group=name,
               text=paste0(
                 "Year: ", year, "<br>",
                 "Value: ", scales::comma(value), "<br>",
                 "Series: ", name, "<br>",
                 "CAGR: ", scales::percent(cagr, accuracy = 0.1)
                 )
               )
           )+
    geom_line()+
    labs(x=NULL, y=NULL, colour=NULL)+
    scale_y_continuous(labels = scales::comma)+
    scale_colour_brewer(palette = "Dark2")

  if(input$origin==TRUE){
    plt <- plt+ 
      expand_limits(y = 0)
  }
  
  plotly::ggplotly(plt, tooltip = "text")   
})


```

### `r renderUI({input$point$text})`

```{r}

DT::renderDT({
  req(input$point$text)
  reactive_base()|>
    filter(slice==input$point$text)|>
    DT::datatable()
})

```

# Comparison with Rich's forecast

## Inputs {.sidebar data-width="300"}

-   We have no alternative forecast to compare with Stokes... so make one up.
-   Rich's forecast is based on LFS and Census NOC *proportions*, and the bottom-line aka aggregate employment forecast.  

### Rich's forecast recipe:

1)  Calculate the NOC proportions for the LFS (annual), and Census 2021.
2)  Calculate the average LFS NOC proportions over the period 2020:2022. (3 years centered on census 2021.)
3)  Re-scale the LFS proportions by multiplying them by the Census proportion and dividing them by the average LFS proportion calculated in 2).
4) Normalize the re-scaled LFS proportions so they sum to one.
5) Create an ensemble forecast of the normalized proportions (TSLS and ETS)
6) Normalize the ensemble proportion forecasts so they sum to one.
7) Apply these normalized proportions to the bottom-line aka aggregate employment forecast. 

### Why?

-  Forecast proportions so aggregation is consistent with the bottom-line forecast.
-  Census to inform levels, LFS for trends.

## Column

### 

```{r}

js2 <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point2', {text: point.text});
  });
}"

plotly::renderPlotly({
 plt <- ggplot(errors, aes(rmse, smape, size=mean_value, colour=mean_value, alpha=mean_value, text=noc_5))+
  geom_point()+
  scale_x_continuous(trans="log10")+
  scale_y_continuous(trans="log10")+
  scale_colour_viridis_c(labels=scales::comma)+
  labs(x="Root Mean Squared Error", y="Symmetric Mean Absolute Percentage Error", colour=NULL)

  plotly::ggplotly(plt, tooltip = "text")|>
     plotly::config(displayModeBar = FALSE)|>
     onRender(js2)
})
```


## Column

### `r renderUI({paste("Employment Share (LFS re-scaled)", input$point2$text, sep=": ")})`


```{r}
plotly::renderPlotly({
  req(input$point2$text)
  plt <- prop_data|>
    filter(noc_5==input$point2$text)|>
    ggplot()+
    geom_point(data=census|>filter(noc_5==input$point2$text), mapping = aes(year, prop, text="Census"))+
    geom_line(aes(year, prop, colour=series))+
    labs(x=NULL, y=NULL, colour=NULL)+
    scale_y_continuous(labels = scales::percent)+
    scale_colour_brewer(palette = "Dark2")
  
  if(input$origin==TRUE){
    plt <- plt+ 
      expand_limits(y = 0)
  }
  
  plotly::ggplotly(plt)
})
```

### `r renderUI({paste("Employment Level", input$point2$text, sep=": ")})`

```{r}
plotly::renderPlotly({
  req(input$point2$text)
  plt <- new_vs_fcast|>
    filter(noc_5==input$point2$text)|>
    group_by(series)|>
    mutate(year=as.numeric(year),
           cagr=(tail(value, n=1)/head(value, n=1))^(1/(max(year)-min(year))) - 1
           )|>
    ggplot()+
    geom_point(data=census|>filter(noc_5==input$point2$text), mapping = aes(year, value, text="Census"))+
    geom_line(aes(year, 
                  value, 
                  colour=series, 
                  group=series, 
                  text=paste0(
                 "Year: ", year, "<br>",
                 "Value: ", scales::comma(value), "<br>",
                 "Series: ", series, "<br>",
                 "CAGR: ", scales::percent(cagr, accuracy = 0.1)
                 )))+
    scale_colour_brewer(palette = "Dark2")+
    labs(x=NULL, y=NULL, colour=NULL)+
    scale_y_continuous(labels = scales::comma)
  
  if(input$origin==TRUE){
    plt <- plt+ 
      expand_limits(y = 0)
  }
  plotly::ggplotly(plt, tooltip = "text")
})
```


