---
title: "4cast comparer"
output:
  flexdashboard::flex_dashboard:
    source_code: https://github.com/bcgov/4cast_viewer_comparison
    orientation: columns
runtime: shiny
resource_files:
- out/joined.rds
- out/cagrs.rds
- out/internal_vs_stokes.rds
- data/lmo64_agg_stokes_mapping.csv
- out/lfs_data.csv
- out/occ_shares.csv
- out/occ_shares_weighted.csv
---

```{r, include=FALSE}
library(tidyverse)
start_year <- today()-years(1) #get rid of -1 once we have this years data
library(here)
library(readxl)
library(janitor)
library(conflicted)
library(htmlwidgets) #for function onRender
library(plotly)
conflicts_prefer(dplyr::filter)
options(scipen = 999)
min_abs_original <- 10
#functions-----------------------
my_dt <- function(tbbl) {
  DT::datatable(tbbl,
                extensions = "Buttons",
                rownames = FALSE,
                filter="top",
                options = list(
                  columnDefs = list(list(className = "dt-center", targets = "_all")),
                  paging = TRUE,
                  scrollX = TRUE,
                  scrollY = TRUE,
                  searching = TRUE,
                  ordering = TRUE,
                  dom = "Btip",
                  buttons = list(
                    list(extend = "csv", filename = "top_5_cor_education"),
                    list(extend = "excel", filename = "top_5_cor_education")
                  ),
                  pageLength = 64,
                  lengthMenu = c(3, 5)
                )
  )
}

cagr_plot <- function(tbbl, xvar, yvar, highlight_industry){
  filtered <- tbbl|>
    filter(industry==highlight_industry)
  plt <- ggplot(cagrs, aes({{  xvar  }}, {{  yvar  }},
                         text=paste0("Industry: ",
                                     industry,
                                     "\n Internal CAGR = ",
                                     scales::percent({{  xvar  }}, accuracy = .1),
                                     "\n Stokes CAGR = ",
                                     scales::percent({{  yvar  }}, accuracy = .1))))+
  geom_abline(slope=1, intercept = 0, colour="white", lwd=1.5)+
  geom_point(data=filtered, size=4, colour="red", alpha=.5)+    
  geom_point(alpha=.5)+
  scale_x_continuous(labels=scales::percent)+
  scale_y_continuous(labels=scales::percent)+
  labs(x="internal forecast",
       y="stokes forecast")
  plotly::ggplotly(plt, tooltip="text")
}
```

```{r}
joined <- read_rds(here("out","joined.rds"))
internal_vs_stokes <- read_rds(here("out","internal_vs_stokes.rds"))
cagrs <- read_rds(here("out","cagrs.rds"))
mapping <- read_csv(here("data","lmo64_agg_stokes_mapping.csv"))
lfs_data <- read_csv(here("out","lfs_data.csv"))
files <- sort(unique(joined$which_file), decreasing = TRUE)
occ_shares <- read_csv(here("out","occ_shares.csv"))
occ_shares_weighted <- read_csv(here("out","occ_shares_weighted.csv"))
```

Comparison with previous cut
=====================================  

Inputs {.sidebar data-width=400}
-------------------------------------

*  This app allows one to visualize what has changed between Stokes cuts, and how employment compares with our internal forecast.
*  The heatmap on this page illustrates the degree of difference between the first and final occupation cut of 2023.
*  Dark colours indicate little difference between cuts:  yellow indicates a (relatively) large difference.
*  You can zoom in by highlighting (click-dragging) a region of the heatmap. 
*  You can **CLICK** on a row in the heatmap to see the underlying data.
*  To zoom back out double click anywhere in the heatmap.


```{r}
selectInput(
  "file",
  "Select an excel file to compare",
  files,
  selected = files[1]
)

which_file <- reactive({
  temp <- joined|>
    filter(which_file==input$file)|>
    pull(joined)
  temp[[1]]
})

renderUI({#need the render UI because the list of choices depends on the excel file
selectInput(
  "variable",
  "Select a variable",
  unique(which_file()$Variable),
  selected = unique(which_file()$Variable)[1]
)
})

renderUI({#need the render UI because the list of choices depends on the excel file
selectInput(
  "geographic_area",
  "Select a geographic area",
  unique(which_file()$`Geographic Area`),
  selected = unique(which_file()$`Geographic Area`)[1]
)
})

reactive_base <- reactive({
  if(length(unique(which_file()$Description))==1 & length(unique(which_file()$Industry))>1){
    temp <- which_file()|>
      rename(slice = Industry)|>
      select(-Description)
  }else{
    temp <- which_file()|>
      rename(slice = Description)|>
      select(-Industry)
  }
  temp|>
    filter(year>=year(today()),
           Variable==input$variable,
           `Geographic Area`==input$geographic_area)|>
    ungroup()|>
    select(-Variable, -`Geographic Area`, -NOC)|>
    group_by(slice)|>
    mutate(original=if_else(min(abs(original_value)) < min_abs_original, "small", "large"),
           change=if_else(original=="small",
                          new_value-original_value,
                          (new_value-original_value)/abs(original_value)*100),
           degree_of_difference=round(log10(abs(change)+1), 3)
          )
})

reactive_tbbl <- reactive({
  reactive_base()|>
  select(slice, year, degree_of_difference)|>
  ungroup()|>
  mutate(slice=fct_reorder(slice, degree_of_difference))|>
  pivot_wider(names_from = "year", values_from = "degree_of_difference")|>
  column_to_rownames("slice")
 })

```
 

Column {data-width=600}
-------------------------------------
    
### 

```{r}

js <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point', {y: point.y});
  });
}"

plotly::renderPlotly({
  req(input$file)
  req(input$variable)
  req(input$geographic_area)
  heatmaply::heatmaply(reactive_tbbl(),
                         dendrogram=FALSE,
                         key.title="Degree of Difference",
                         fontsize_row = 8,
                         showticklabels = c(TRUE, FALSE))|>
    plotly::config(displayModeBar = FALSE)|>
    onRender(js)
})
```

Column {data-width=400}
-------------------------------------

### Difference for `r renderUI({input$variable})`

```{r}
plotly::renderPlotly({
  req(input$point)
  plt <- reactive_base()|>
    filter(slice==rev(rownames(reactive_tbbl()))[input$point$y])|>
    mutate(new_10_cagr=((tail(new_value, n=1)-head(new_value, n=1)+abs(head(new_value, n=1)))/(abs(head(new_value, n=1))))^(.1)-1,
           original_10_cagr=((tail(original_value, n=1)-head(original_value, n=1)+abs(head(original_value, n=1)))/(abs(head(original_value, n=1))))^(.1)-1,
           new_10_cagr=scales::percent(new_10_cagr, accuracy=.1),
           original_10_cagr=scales::percent(original_10_cagr, accuracy=.1)
          )|>
    pivot_longer(cols = c(new_value, original_value))|>
    mutate(year=as.numeric(year))|>
    ggplot(aes(year, 
               value, 
               colour=name,
               text=paste0(
                 "Original 10 year CAGR: ",
                 original_10_cagr,
                 "\n New 10 year CAGR: ",
                 new_10_cagr)))+
    geom_line()+
    labs(title=rev(rownames(reactive_tbbl()))[input$point$y])
  plotly::ggplotly(plt)
})
```

### The data:

```{r}
DT::renderDT({ 
  req(input$point)
  reactive_base()|>
    filter(slice==rev(rownames(reactive_tbbl()))[input$point$y])|>
    select(year, new_value, original_value)|>
    DT::datatable(rownames = FALSE)
})
```

Comparison with internal employment forecast
=====================================  

Column
-------------------------------------
    
### Degree of difference between internal and stokes:  **CLICK** on row to choose industry
    
```{r}

js2 <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point2', {y: point.y});
  });
}"

temp <- internal_vs_stokes|>
  unnest(data)|>
  mutate(change=(stokes_cut/internal-1)*100,
         degree_of_difference=round(log10(abs(change)+1), 3))|>
  select(industry, year, degree_of_difference)|>
  filter(year>2023)|>
  pivot_wider(id_cols = industry, names_from = "year", values_from = "degree_of_difference")|>
  column_to_rownames("industry")

  heatmaply::heatmaply(temp, dendrogram=FALSE, key.title="Degree of Difference", showticklabels = c(TRUE, FALSE))|>
  onRender(js2)
```
 
Column
------------------------------------- 
 
### First five year CAGRs 
    
```{r}
plotly::renderPlotly({
req(input$point2)  
cagr_plot(cagrs, internal_cagr_ffy, stokes_cagr_ffy, rev(rownames(temp))[input$point2$y])
})  
``` 


### Second five year CAGRs
    
```{r}
plotly::renderPlotly({
req(input$point2)    
cagr_plot(cagrs, internal_cagr_sfy, stokes_cagr_sfy, rev(rownames(temp))[input$point2$y])
})  
```

Column
------------------------------------- 

### Ten year CAGRs

```{r}
plotly::renderPlotly({
req(input$point2)    
cagr_plot(cagrs, internal_cagr_ty, stokes_cagr_ty, rev(rownames(temp))[input$point2$y])
})  
```

### `r renderUI({rev(rownames(temp))[input$point2$y]})` relative to LFS Data


```{r}
reactive_tbbl2 <- reactive({
 internal_vs_stokes|>
    unnest(data)|>
    mutate(date=ymd(paste(year,"06","15", sep="-")))|>
    filter(date>start_year)|>
    select(-year)|>
    pivot_longer(cols=c(stokes_cut, internal), names_to = "series", values_to = "value")|>
    full_join(lfs_data)|>
    filter(industry==rev(rownames(temp))[input$point2$y])
})

reactive_lfs_last_year <- reactive({
  lfs_data|>
    filter(industry==rev(rownames(temp))[input$point2$y])|>
    group_by(industry)|>
    mutate(value=zoo::rollmean(value, 12, align = "right", fill=NA),
           series="12 month rolling mean of LFS")
})

plotly::renderPlotly({
  req(input$point2) 
  plt <- ggplot(reactive_tbbl2(), aes(date, value, colour=series))+
    geom_vline(xintercept = as.numeric(floor_date(today(), unit="year")-months(1)), colour="white", lwd=1)+
    geom_line(alpha=.5)+
    geom_line(data=reactive_lfs_last_year())+
    scale_y_continuous(labels = scales::comma)+
    scale_colour_brewer(palette = "Dark2")+
    labs(x=NULL,
         y=NULL,
         colour=NULL)
  plotly::ggplotly(plt)|>
    plotly::layout(legend = list(orientation = "h", x = 0.4, y = -0.2))
})
```


Mapping from LMO industries to Stokes industries
===================================== 

```{r}
my_dt(mapping)
```

Occupation Shares
=====================================================

Inputs {.sidebar}
-------------------------------------

* The scatterplot to the right gives an indication of a "jump" between the historic shares and the forecast shares
* The white diagonal line indicates no jump, and the further you are from the line the larger the jump in share of total employment.
* You can **click** on a point to see the underlying time series of employment shares.


Column {data-width=650}
-----------------------------------------------------------------------

### Shares of total employment: LFS vs. Stokes

```{r}

js3 <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point3', {text: point.text});
  });
}"
renderPlotly({
  plt <- ggplot(occ_shares_weighted, aes(lfs,
                       stokes,
                       text=Description))+
  geom_abline(slope = 1, intercept = 0, colour="white", lwd=2)+
  geom_point(alpha=.2)+
  scale_x_continuous(trans="log10", labels = scales::percent)+
  scale_y_continuous(trans="log10", labels = scales::percent)+
  labs(x="Shares based on the Labour Force Survey",
       y="Shares based on Stokes forecast"
       )
  ggplotly(plt, tooltip = "text")|>
    onRender(js3)
})
```


Column {data-width=350}
-----------------------------------------------------------------------

### `r renderUI({input$point3$text})`

```{r}
renderPlotly({
  req(input$point3)
  plt <- occ_shares|>
    filter(Description==input$point3$text)|>
    ggplot(aes(year, 
               share, 
               colour=sample,
              ))+
    geom_line()+
    scale_y_continuous(labels=scales::percent)+
    coord_cartesian(ylim = c(0, NA))+
    labs(colour=NULL)
  ggplotly(plt)
})
```
