---
title: "4cast comparer"
output:
  flexdashboard::flex_dashboard:
    source_code: https://github.com/bcgov/4cast_viewer_comparison
    orientation: columns
runtime: shiny
---

```{r, include=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(conflicted)
library(htmlwidgets) #for function onRender
library(plotly)
conflicts_prefer(dplyr::filter)
options(scipen = 999)
#functions-----------------------
my_dt <- function(tbbl) {
  DT::datatable(tbbl,
                extensions = "Buttons",
                rownames = FALSE,
                filter="top",
                options = list(
                  columnDefs = list(list(className = "dt-center", targets = "_all")),
                  paging = TRUE,
                  scrollX = TRUE,
                  scrollY = TRUE,
                  searching = TRUE,
                  ordering = TRUE,
                  dom = "Btip",
                  buttons = list(
                    list(extend = "csv", filename = "top_5_cor_education"),
                    list(extend = "excel", filename = "top_5_cor_education")
                  ),
                  pageLength = 64,
                  lengthMenu = c(3, 5)
                )
  )
}
symmetric_change <- function(old, new) {
  denom <- (old + new) / 2
  result <- ifelse(denom == 0 & new != old, 200 * sign(new - old),
                   ifelse(denom == 0, 0, (new - old) / denom * 100))
  return(result)
}
percent_level_change <- function(tbbl){
  old <- sum(tbbl$original_value)
  new <- sum(tbbl$new_value)
  symmetric_change(old, new)
}

how_big <- function(tbbl){
  mean(tbbl$new_value)
}

percent_cagr_change <- function(tbbl){
  tbbl <- tbbl|>
    mutate(year=as.numeric(year))|>
    arrange(year)
  new <- (last(tbbl$new_value)/first(tbbl$new_value))^(1/(last(tbbl$year)-first(tbbl$year)))
  old <-(last(tbbl$original_value)/first(tbbl$original_value))^(1/(last(tbbl$year)-first(tbbl$year)))
  symmetric_change(old, new)
}
```

```{r}
new_vs_old <- read_rds(here("out","new_vs_old.rds"))
files <- sort(unique(new_vs_old$which_file), decreasing = TRUE)
new_vs_fcast <- read_rds(here("out","new_vs_fcast.rds"))
errors <- read_rds(here("out","errors.rds"))
```

# Comparison with previous cut

## Inputs {.sidebar data-width="300"}

-   This app allows for two comparisons:

1)  between two Stokes occupational forecasts.
2)  between the current Stokes occupational employment forecast and historic LFS data.

-   For the chart to the immediate left, we create 3 summary statistics for the selected data

1)  The percentage difference in the mean level (across 10 years)
2)  The percentage difference in 10 year growth rates
3)  The size of the "slice" of data (i.e. the size of industry or occupation.)

-   For points close the origin, Stokes has not made changes that affect either the level or the growth rate.
-   For points far away from the origin, there has been a substantial change (in percentage terms.)

```{r}
selectInput(
  "file",
  "Select an excel file to compare",
  files,
  selected = files[4]
)

which_file <- reactive({
  temp <- new_vs_old|>
    filter(which_file==input$file)|>
    pull(joined)
  temp[[1]]
})

renderUI({#need the render UI because the list of choices depends on the excel file
selectInput(
  "variable",
  "Select a variable",
  unique(which_file()$Variable),
  selected = unique(which_file()$Variable)[1]
)
})

renderUI({#need the render UI because the list of choices depends on the excel file
selectInput(
  "geographic_area",
  "Select a geographic area",
  unique(which_file()$`Geographic Area`),
  selected = unique(which_file()$`Geographic Area`)[1]
)
})

checkboxInput("origin", "Are you Sazid? (should y axis include the origin)", value = FALSE)

#reactive elements------------------------------

reactive_base <- reactive({
  if(length(unique(which_file()$Description))==1 & length(unique(which_file()$Industry))>1){
    temp <- which_file()|>
      filter(Industry!="All industries")|>
      rename(slice = Industry)|>
      select(-Description)
  }else{
    temp <- which_file()|>
      filter(Description!="Total")|>
      rename(slice = Description)|>
      select(-Industry)
  }
  temp|>
    filter(year>=year(today()),
           Variable==input$variable,
           `Geographic Area`==input$geographic_area)|>
    ungroup()|>
    select(-Variable, -`Geographic Area`, -NOC)|>
    group_by(slice)
})

reactive_tbbl <- reactive({
  reactive_base()|>
    mutate(year=as.numeric(year))|>
    filter(year >= max(year)-10)|>
  group_by(slice)|>
  nest()|>
  mutate(percent_level_change=map_dbl(data, percent_level_change),
         percent_cagr_change=map_dbl(data, percent_cagr_change),
         mean_level=map_dbl(data, how_big))|>
  unnest(data)
})
```

## Column

### `r renderUI({paste(input$geographic_area, input$variable, sep=": ")})`

```{r}

js <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point', {text: point.text});
  });
}"

plotly::renderPlotly({
  req(input$variable)
  req(input$geographic_area)
  plt <- reactive_tbbl()|>
  select(slice, starts_with("percent"), mean_level)|>
  distinct()|>
  ggplot(aes(percent_level_change,
             percent_cagr_change,
             size=mean_level,
             alpha=mean_level,
             colour=mean_level,
             text=slice))+
  geom_vline(xintercept = 0, colour="white", lwd=2)+
  geom_hline(yintercept = 0, colour="white", lwd=2)+ 
  geom_point()+
  scale_colour_viridis_c()+
  labs(x="Level Change (%)",
       y="Growth Change (%)",
       colour="Mean Level (new)")


plotly::ggplotly(plt, tooltip = "text")|>
     plotly::config(displayModeBar = FALSE)|>
     onRender(js)
})
```

## Column

### `r renderUI({input$point$text})`

```{r}
plotly::renderPlotly({
  req(input$point)
  plt <- reactive_base()|>
    filter(slice==input$point$text)|>
    pivot_longer(cols = c(new_value, original_value))|>
    mutate(year=as.numeric(year))|>
    ggplot(aes(year,
               value,
               colour=name,
               text=slice))+
    geom_line()+
    scale_y_continuous(labels=scales::comma)+
    labs(x=NULL,y=NULL, colour=NULL)
  
  if(input$origin==TRUE){
    plt <- plt+ 
      expand_limits(y = 0)
  }

  plotly::ggplotly(plt)
})
```

### `r renderUI({input$point$text})`

```{r}
DT::renderDT({
  req(input$point)
  reactive_tbbl()|>
    filter(slice==input$point$text)|>
    ungroup()|>
    select(-slice)|>
    mutate(across(everything(), round, 2))|>
    DT::datatable(rownames = FALSE)
})
```

# Comparison with ensemble forecast

## Inputs {.sidebar data-width="300"}

-   On this page we compare the most recent Stokes forecast $S_t$ with an ensemble forecast $E_t$ based on LFS data since 2000.
-   The ensemble forecast is based on 50/50 mixture of two component forecasts:

1)  TSLM: Linear regression of employment on time.
2)  ETS: [Exponential Smoothing](https://otexts.com/fpp3/expsmooth.html)

-   We use two metrics to compare the stokes forecast with the ensemble forecast, which are displayed in the scatterplot to the right.

$${\displaystyle \operatorname {RMSE} ={\sqrt {{\frac {1}{n}}\sum _{t=1}^{n}(S_{t}-E_{t})^{2}}}}$$

$${\displaystyle {\text{SMAPE}}={\frac {1}{n}}\sum _{t=1}^{n}{\frac {|S_{t}-E_{t}|}{|E_{t}|+|S_{t}|}}}$$

- Bottom left forecasts are similar, top right the forecasts differ.


## Column

### 

```{r}

js2 <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point2', {text: point.text});
  });
}"

plotly::renderPlotly({
 plt <- ggplot(errors, aes(rmse, smape, size=mean_value, colour=mean_value, alpha=mean_value, text=noc_5))+
  geom_point()+
  scale_x_continuous(trans="log10")+
  scale_y_continuous(trans="log10")+
  scale_colour_viridis_c()

  plotly::ggplotly(plt, tooltip = "text")|>
     plotly::config(displayModeBar = FALSE)|>
     onRender(js2)
})
```


## Column

###  `r renderUI({input$point2$text})`

```{r}
plotly::renderPlotly({
  req(input$point2$text)
  plt <- new_vs_fcast|>
    filter(noc_5==input$point2$text)|>
    ggplot(aes(syear, .mean, colour=.model))+
    geom_line()+
    labs(x=NULL, y=NULL, colour=NULL)+
    scale_y_continuous(labels = scales::comma)
  plotly::ggplotly(plt)
})
```

### `r renderUI({input$point2$text})`

```{r}
DT::renderDT({
  req(input$point2$text)
  new_vs_fcast|>
    filter(noc_5==input$point2$text)|>
    ungroup()|>
    select(year=syear, value=.mean, series=.model)|>
    mutate(value=round(value))|>
    arrange(year)|>
    DT::datatable(rownames = FALSE)
})
```


